The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-03-13 analysis/paper/paper.qmd
Explore topic model

```{r}
#| label: extract-pdf-text
#| echo: false
emu_theses$text <- convert_pdf_text(pdf_paths)
```

@@

emu_theses$text_clean <- emu_theses$text |>
  clean_basic() |>
  str_remove_all(find_meta_stopwords(emu_theses)) |> 
  str_remove_all(subject_stopwords()) |> 
  str_remove_all(document_stopwords())
```

@@

urbanism_stopwords <- urbanism_stopwords() 
thesis_stopwords <- thesis_stopwords()

emu_theses_words <- emu_theses |> 
  select(title, text_clean) |> 
  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, seperate all words
  anti_join(stop_words, by = "word") |> 
  anti_join(urbanism_stopwords, by = "word") |> 
  anti_join(thesis_stopwords, by = "word")
```

@@

>>>>

```{r}
#| label: extract-pdf-text
#| echo: false
#| cache: true

emu_theses$text <- convert_pdf_text(pdf_paths)
```

@@

emu_theses$text_clean <- emu_theses$text |> 
  clean_basic() |> 
  str_remove_all(find_meta_stopwords(emu_theses) )  
```

@@

stop_words_custom <- emu_theses %>% 
  unnest_tokens(output = word, input = title) %>%   ### add words from the title
  select(word) %>% 
  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
                               "source", "author"))) %>% 
  # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
  rbind(., data.frame(word = c("hab", "km")))

emu_theses_words <- emu_theses %>%
  mutate(author = paste(first_name, last_name)) %>% 
  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
  unnest_tokens(output = word, input = text) %>%   # remove punctuation, convert to lowercase, seperate all words
  anti_join(stop_words, by = "word") %>%  # remove stop words
  anti_join(stop_words_custom, by = "word")
```

@@

## Topic model

Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].

```{r}
# The topicmodels package requires a document-term matrix as input
emu_theses_dtm <- emu_theses_words |> 
  count(title, word, sort = TRUE) |> 
  cast_dtm(title, word, n)

emu_theses_dtm

# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
# In LDA, each document is a mixture of topics and each topic is a mixture of words
emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
emu_theses_lda

# Tidy model with beta statistic
emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
emu_theses_topics

# The top 10 words that are the most common in each topic
emu_theses_top_terms <- emu_theses_topics |> 
  group_by(topic) |> 
  top_n(20, beta) |> 
  ungroup() |> 
  arrange(topic, -beta)

emu_theses_top_terms |> 
  mutate(term = reorder(term, beta)) |> 
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

# 
beta_spread <- emu_theses_topics |> 
  mutate(topic = paste0("topic", topic)) |> 
  spread(topic, beta) |> 
  filter(topic1 > 0.001, topic2 > .001) |> 
  mutate(log_ratio = log2(topic2/topic1))

beta_spread

beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  slice_max(abs(log_ratio), n = 20) %>% 
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(log_ratio, term)) +
  geom_col() +
  labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)

emu_theses_documents <- tidy(emu_theses_lda, matrix = "gamma")
emu_theses_documents

emu_theses_documents |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip()

emu_theses_documents |> 
  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)

emu_theses_documents |> 
  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
  geom_col()
```
